---
title: "Reproducible Research: Peer Assessment 1"
author: Bora Aslan
date: August 19, 2018

---

## Loading and preprocessing the data
Loading and some process on the data
```{r}
unzip("activity.zip")
myData <- read.csv("activity.csv",sep=",",head=TRUE,colClasses=c("integer","Date","integer"))
myDataFinal <- myData[!is.na(myData$steps),]
head(myDataFinal)
```

## What is mean total number of steps taken per day?
The following operations are done in order
1, Split the data
2. Plot
3. Find the mean
4. Find the median

```{r}
library(plyr)
salesPerDay<- sapply(split(myDataFinal$steps,myDataFinal$date),sum,na.rm=TRUE)
hist(salesPerDay)
meanSales <- mean(salesPerDay)
medianSales <- median(salesPerDay)
```
The mean of sales per day is **`r meanSales`** 
The median of sales is **`r medianSales`.** 

## What is the average daily activity pattern?
The following operations are done in order
1. Split in intervals
2. Plot

```{r}
IntervalSteps <- ddply(myDataFinal,~interval,summarise, mean_steps=mean(steps))
plot(IntervalSteps$interval, IntervalSteps$mean_steps, ylab = "Average Number of Steps Taken", xlab = "Interval", type = "l")

```

## Imputing missing values
1. Impute the missing values 
2. Plot 
3. Find the mean again
4. Find the median again

```{r}
myDataFinal <- myData[!is.na(myData$steps),]
nrow(na.omit(myDataFinal))
dataCheck<-myDataFinal
for(i in 1:nrow(myDataFinal)){
    if(is.na(myDataFinal$steps[i])){
        dataCheck$steps[i]<- salesPerDay[(i %% 288)+1]
    }
}
```


```{r}
salesPerDayNew <- sapply(split(dataCheck$steps,dataCheck$date),sum,na.rm=T)
hist(salesPerDayNew)

meanSalesNew <- mean(salesPerDayNew)
medianSalesNew <- median(salesPerDayNew)

```

After the new computations, the updated mean and median are as follows
The new mean of number of steps is **`r meanSalesNew`** .  
The new mediane is **`r medianSalesNew**.

### The 5-minute interval that, on average, contains the maximum number of steps?

```{r}
dataInterval<- ddply(dataCheck,~interval,summarise, mean_steps=mean(steps))
maxInterval <- max(dataInterval$mean_steps)
theInterval <- max(dataCheck[dataInterval$mean_steps == maxInterval, 1])
```
The maximum number of steps is ***`r maxInterval`***
The 5-minute interval is ***`r theInterval`***

## Are there differences in activity patterns between weekdays and weekends?

```{r}
library(lattice)
weekDAY <- weekdays(as.Date(dataCheck$date))
dataWeekDAY <- transform(dataCheck, day = weekDAY)
dataWeekDAY$wk <- ifelse(dataWeekDAY$day %in% c("Saturday", "Sunday"), "weekend", 
    "weekday")
average_week <- ddply(dataWeekDAY, .(interval, wk), summarise, steps = mean(steps))

xyplot(steps ~ interval | wk, data = average_week, layout = c(1, 2), type = "l")
```


### We can see from the graph the comparison between the people walking on weekdays and weekends


### Use the Knit option in RStudio to create the Html file.





